Deciphering the Trends: A Comprehensive Analysis of US Birth Rates and Educational Attainment from 1980 to 2020

### 2024-02-14 outputs/paper/paper.qmd
new references

The most notable decrease was seen among teenagers, where birth rates fell from 53 per 1,000 in 1980 to 15.3 in 2020. This dramatic decline not only reflects the success of public health initiatives aimed at reducing teenage pregnancies but also suggests a shift in the social fabric, with young individuals increasingly prioritizing education and career over early family formation [@cnbc]. The decline in birth rates across all age groups from 1980 to 2020 signifies a profound transformation in societal norms and individual life choices.

The decline in birth rates among women in their 20s, while not as sharp as that of teenagers, nevertheless indicates a substantial shift in the way younger women view motherhood. The change in birth rates from 115.1 births per 1,000 women in the 20–24 age group in 1980 to 62.8 in 2020, and from 112.9 to 90.0 in the 25–29 age group in the same period, reflects broader trends in society that prioritize the pursuit of higher education, career advancement, and personal development over having children at a young age [@cnbc].

@@

Future research should aim to investigate the relationship between financial stability and fertility decisions by correlating trends in birth rates with economic data. Questions like whether having a higher income causes one to have more children or if people are delaying having children because of financial concerns should be thoroughly examined. 
Other possibly influences on fertility decisions should also be examined at a more social level, not just economic. Changing policy and attitudes will constanly be a force in the number of children being born. Looking into how shock changes in societies structure would be a interesting topic of analysis. An example of a topic of research would be the impact of the Roe v. Wade political changes across statees in the US.
Medical advancements are another changing field, examining the implications of an aging population and the social and economic costs of altering fertility patterns could offer insightful information to policymakers and society at large.

>>>>

The most notable decrease was seen among teenagers, where birth rates fell from 53 per 1,000 in 1980 to 15.3 in 2020. This dramatic decline not only reflects the success of public health initiatives aimed at reducing teenage pregnancies but also suggests a shift in the social fabric, with young individuals increasingly prioritizing education and career over early family formation [@manlove]. The decline in birth rates across all age groups from 1980 to 2020 signifies a profound transformation in societal norms and individual life choices.

The decline in birth rates among women in their 20s, while not as sharp as that of teenagers, nevertheless indicates a substantial shift in the way younger women view motherhood. The change in birth rates from 115.1 births per 1,000 women in the 20–24 age group in 1980 to 62.8 in 2020, and from 112.9 to 90.0 in the 25–29 age group in the same period, reflects broader trends in society that prioritize the pursuit of higher education, career advancement, and personal development over having children at a young age [@manlove].

@@

Future research should aim to investigate the relationship between financial stability and fertility decisions by correlating trends in birth rates with economic data. Questions like whether having a higher income causes one to have more children or if people are delaying having children because of financial concerns should be thoroughly examined. 
Other possibly influences on fertility decisions should also be examined at a more social level, not just economic. Changing policy and attitudes will constanly be a force in the number of children being born. Looking into how shock changes in societies structure would be a interesting topic of analysis. Recent repeals of Roe v. Wade is an example of such shock changes that will greatly impact fertility levels in the US (@roevwade).
Medical advancements are another changing field, examining the implications of an aging population and the social and economic costs of altering fertility patterns could offer insightful information to policymakers and society at large.
